---
title: "Example Air Visit Data Format"
author: "Singh, Gurjeet"
date: "February 04, 2018"
output:
  pdf_document:
   # toc: true
   # toc_depth: 3  
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(tibble)
library(fpp)
```


## 1. Example: Source Data

I am importing and reformatting the source data to a useful format for time series.\newline

```{r Import & Format Data}
visit.data.df <- read_csv("Singh_Gurjeet_Example_air_visit_data.csv")
visit.df <-   visit.data.df %>% 
              spread(air_store_id, visitors) %>% 
              separate(visit_date, c("Year","Month","Day"))

```

## 2. Example: Define Time Series

I am converting the data into a time series and defining the time series object.\newline
```{r Define Time Series}
newdata=as.ts(visit.df[, -c(1,2,3)], start=c(2016,1,1), frequency=365) 
```

## 3. Example: Simple Imputation

I am performing a quick and simple imputation for demonstration purposes.\newline
```{r Silly Imputation Example}

for(i in 1:ncol(newdata)){
    newdata[is.na(newdata[,i]), i] <- mean(newdata[,i], na.rm = TRUE)
}
```

## 4. Example: Simple Forecasting (ETS)

I am using the for loop to forecast for each store id for 39 days.\newline
```{r Simple EST Forecast Example}
myf2=round(forecast(ets(newdata[,1]),39)$mean,0)
#simple ETS
for (i in 2:829) {
  myf= round(forecast(ets(newdata[,i]),39)$mean,0)
  myf2=cbind(myf,myf2)
}
colnames(myf2)=colnames(newdata)
```


## 5. Example: Final Submission Data 

I am converting the forecast result to a data frame with the row id. Next, I format the data into a format of a submission file. Lastly, I save the file in a CSV format.\newline

```{r Format & Export Data}
data.result <- rownames_to_column(as.tibble(myf2),var = "RowID")
final_submission <- as.tibble(data.result %>% 
                              mutate(Date = seq(from = as.Date("2017-04-23"), 
                                                to = as.Date("2017-05-31"), 
                                                by= 'day')) %>% 
                              gather("StoreID","visitors",
                                     2:ncol(data.result)) %>% 
                              unite("id",StoreID,Date, sep = "_")
                             )[,2:3]
write_csv(final_submission, "Singh_Gurjeet_Example_final_submission.csv")
```

